Efficient N-Dimensional surface estimation using Crofton formula and run-length encoding

### 2012-02-06 article/Article.tex
add missing computation of lMin and lMax
ajoute equations sommes, ajoute schema poids des directions, biblio

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  hyperref should be the last package to be loaded.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}
% to be able to use options in graphics
\usepackage{graphicx}
% for pseudo code
\usepackage{listings}
% subfigures
\usepackage{subfigure}
\usepackage{pseudocode}

@@

@@

The Crofton method is an alternative method that allows measuring the perimeter of 2D objects,
the surface area of 3D objects, and more generally the $(d-1)$-dimensional measure of $d$-dimensional 
object boundary. 
\commentaire{Ajouter un petit argumentaire sur convergence des mesures avec Crofton.}

@@

% Crofton formula
Perimeter measure of 2D objects, surface area measure of 3D objects 
and more generally $(d-1)$-surface measure of $d$-dimensional objects 
can be expressed in a unified formalism by using the Crofton formula.
This formula consists in integrating the intercept number of the object with
lines of various orientation and positions. Its expression in
the general form is:
\begin{eqnarray}
S^{(d-1)}(X) & = & \frac{d \cdot v_d}{\pi}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
\end{eqnarray}
where $X$ is the structure of interest, $v_d$ is the volume of the $d$-dimensional ball,
$\mathcal{L}^d$ is the set of all lines in the $d$-dimensional space. 
The above integral is normalised such that the mass of lines hitting the unit ball 
equals $v_{d-1}$, the measure of the unit ball projection on a $(d-1)$-dimensional plane.

@@

The Crofton formula can be easily applied to discrete binary images \cite{Lang2001, Legland2007}. 
The integral over lines can be decomposed into an integral over a finite set of directions
and an integral over lines parallel to a given direction.
The number of connected components in the intersections of the structure with a discrete line
is obtained by counting the number of intercepts with its boundary.
\commentaire{ 
Ajouter une formule avec des sommes a la place des integrales (David)}

\commentaire{dire comment on calcule les poids (1) des espaces entre les droites et (2) des poids associes aux directions}

\commentaire{Cas des directions principales ($d$ en $nd$) et de toutes les directions (4 en 2D et 13 en 3D)}

@@

The intercepts count, required to use the method described earlier, can be efficiently computed by using the run-length
encoding of the binary objects.

@@

\begin{figure}[htbp]
\centering
\small
\begin{pseudocode}[framebox]{countIntercept}{O}
\COMMENT{intercept count} \\
ic \GETS \emptyset \\
\COMMENT{the offset of the neighbors on the first axis} \\
xno \GETS offset(0) \\
xno[0] \GETS 1 \\
\FOREACH ls \in O \DO
\BEGIN
  \COMMENT{2 intercepts per line on the first axis} \\
  ic[xno] \GETS ic[xno] + 2 \cdot length(ls) \\
  \FOREACH nls \in neighbors(ls) \DO
  \BEGIN
    \COMMENT{straight and diagonal offsets of the neighbors} \\
    no \GETS offset(nls) \\
    dno \GETS offset(nls) \\
    dno[0] \GETS 1 \\
    \IF isEmpty(nls) \THEN
    \BEGIN
      \COMMENT{all the lines are on a contour} \\
      \FOREACH l \in ls \DO
      \BEGIN
	ic[no] \GETS ic[no] + length(l)\\
	ic[dno] \GETS ic[dno] + 2 \cdot length(l)
      \END
    \END
    \ELSE
    \BEGIN
      \COMMENT{iterate over the lines in $ls$ and the empty lines in $nls$} \\
      il \GETS iterator(ls) \\
      inl \GETS iterator(nls) \\
      nMin \GETS -\infty \\
      nMax \GETS pos(inl) - 1 \\
      \WHILE isNotAtEnd(il) \DO
      \BEGIN
        \COMMENT{measure the intersection of these two lines} \\
	ic[no] \GETS ic[no] + max( 0, min( lMax, nMax ) - max( lMin, nMin ) + 1 ) \\
	ic[dno] \GETS ic[dno] + max( 0, min( lMax, nMax+1 ) - max( lMin, nMin+1 ) + 1 )\\
	ic[dno] \GETS ic[dno] + max( 0, min( lMax, nMax-1 ) - max( lMin, nMin-1 ) + 1 )\\
        \COMMENT{and move to the next line, either in $ls$ or in $nls$} \\
	\IF nMax \leq lMax \THEN
	\BEGIN
          nMin \GETS pos(inl) + length(inl) \\
          next(inl) \\
          \IF isNotAtEnd(inl) \DO
            nMax \GETS pos(inl) - 1
          \ELSE
            nMax \GETS \infty
	\END
	\ELSE
          next(il)
      \END
    \END
  \END
\END
\end{pseudocode}
\caption{\label{interceptCountAlgo}The algorithm used to count the intercepts}
\end{figure}

@@

The roundness is defined the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
The equivalent $d$-perimeter is defined as the perimeter of the $d$-ball with same volume $V$.
\begin{eqnarray}
roundness             & = & \frac{P_{eq}}{P} \\
P_{eq}                & = & \frac{d \cdot V_{eq}}{R_{eq}} \\
R_{eq}                & = & \sqrt[\displaystyle d]{\frac{V\cdot\Gamma(\frac{d+1}{2})}{\pi^\frac{d}{2}}} \\
\Gamma(\frac{d+1}{2}) & = & \begin{cases}
                               \displaystyle \frac{d}{2}!  &\text{if}~ d ~\text{is even} \\
                               \displaystyle \frac{\sqrt{\pi} \cdot d!!}{2^{\frac{d+1}{2}}}  &\text{if}~ d ~\text{is odd} \\
                            \end{cases} \\
d!!                   & = & \begin{cases}
                                1 &\text{if}~ d \leq 1 \\
                                d \cdot (d-2)!!  & \text{otherwise} \\
                             \end{cases}
\end{eqnarray}

@@

\subsection{Shape discretization}

A 2D or 3D binary image can be considered as a subset of a rectangular grid $L_d$ where
d = 2,3. Such a grid can be written as
\begin{eqnarray}
L_d = \Delta_{1} \mathbb{Z} \times \ldots \times \Delta_{1} \mathbb{Z},
\end{eqnarray}

where the $d$-uple $(\Delta_{1},\ldots,\Delta_{d})$ defines the pixel
or voxel size. A digitization scheme represents how a continuous
structure $X$ is digitized into a binary image $X_\text{d}$. 
According to the
covering scheme a grid point $x$ belongs to $X_\text{d}$ if the grid
cell centered on $x$ hits $X$ (See Fig.~\ref{fig:DiskDiscretization}).

>>>>

% to be able to use options in graphics
\usepackage{graphicx}
% for pseudo code
\usepackage{listings}
% subfigures
\usepackage{subfigure}
\usepackage{pseudocode}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  hyperref should be the last package to be loaded.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}

@@

The Crofton method is an alternative method that allows estimating the perimeter of 2D objects,
the surface area of 3D objects, and more generally the $(d-1)$-dimensional measure of $d$-dimensional 
object boundary. 
It is based on counting intercept number of the object boundary with a set of isotropic test lines, 
and provide an unbiased estimate of the actual perimeter or surface.

@@

% Crofton formula
Perimeter measure of 2D objects, surface area measure of 3D objects 
and more generally $(d-1)$-surface measure of $d$-dimensional objects 
can be expressed in a unified formalism by using the Crofton formula.
This formula consists in integrating the intercept number of the object with
lines of various orientation and positions. Its expression in
the general form is:
\begin{eqnarray}
S^{(d-1)}(X) & = & \frac{d \cdot v_d}{v_{d-1}}\int_{\mathcal{L}^d}\chi\left(X\cap L\right) dL
\end{eqnarray}
where $X$ is the structure of interest, $v_d$ is the volume of the $d$-dimensional ball,
$\mathcal{L}^d$ is the set of all lines in the $d$-dimensional space. 
The above integral is normalised such that the mass of lines hitting the unit ball 
equals $v_{d-1}$, the measure of the unit ball projection on a $(d-1)$-dimensional plane.

@@

The Crofton formula can be easily applied to discrete binary images \cite{Lang2001, Legland2007}. 
The integral over lines can be decomposed into an integral over a finite set of directions
and an integral over all the lines parallel to a given direction.

For the planar case, the perimeter can be estimated by considering horizontal and vertical lines, 
i.e.\ two directions. An alternative is to use the diagonals, resulting in four directions. 
Perimeter estimate is then written as:
\begin{eqnarray}
P(X) & \simeq & \pi \sum_{k}\frac{c_{k}}{\lambda_{k}}\chi \left( X \cap L_{k} \right)
\end{eqnarray}
where $c_k$ is the discretization weight associated to direction $k$, 
$\lambda_k$ is the density of discrete lines in direction $k$, 
and $L_k$ is the set of all discrete lines in direction $k$.

The line density $\lambda_k$ may vary according to the directions, 
due to image resolution or to the use of diagonals. 
It is computed as the ratio of the distance between two neighbor pixels with
the area associated to a pixel.
When only horizontal and vertical directions are used, associated weights $c_k$ equal $1/2$.
When diagonals are used, weights are obtained by projecting direction vectors on the unit circle, 
and by computing the relative fraction of circle associated to each direction 
(Fig.~\ref{fig:DirectionWeights2D3D}).

% Compute direction weights
\begin{figure}[!htb]
\begin{center}
\includegraphics[width=8cm]{images/directionsWeights2D}
\end{center}
\caption{Computation of direction weights for rectangular pixel grid. 
Black dots correspond to pixel centers. 
Angular regions are computed between direction bisectors.}
\label{fig:DirectionWeights2D3D}
\end{figure}

In a similar way, surface area of $3D$ objects may be estimated using following approximation:
\begin{eqnarray}
S(X) & \simeq & 4\sum_{k}\frac{c_{k}}{\lambda_{k}}\chi \left( X \cap L_{k} \right)
\end{eqnarray}
where $L_k$ is the set of $3D$ discrete lines parallel to direction $k$, 
$c_k$ the discretization weight associated to $3D$ direction $k$, 
and $\lambda_k$ the density of discrete lines in direction $k$.

Line densities are computed as for the planar case, by dividing distance of neighbor
voxels in direction $k$ by the volume associated to a single voxel.
When surface area is estimated by taking into account the three main direction in images, 
associated weights $c_k$ equal $1/3$. If more directions are considered, the strategy of
\cite{Ohser2000} was used: each direction was projected on the unit sphere, 
and used as a germ for computing Voronoï diagram on the unit sphere. 
The relative surface area of each spherical domain was used as weight for the
corresponding direction.

@@

The number of connected components in the intersections of the structure with a discrete line
is obtained by counting the number of intercepts with its boundary.
This intercepts count can be efficiently computed by using the run-length
encoding of the binary objects.

@@

\begin{figure}[htbp]
\centering
\small
\begin{pseudocode}[framebox]{countIntercept}{O}
\COMMENT{intercept count} \\
ic \GETS \emptyset \\
\COMMENT{the offset of the neighbors on the first axis} \\
xno \GETS offset(0) \\
xno[0] \GETS 1 \\
\FOREACH ls \in O \DO
\BEGIN
  \COMMENT{2 intercepts per line on the first axis} \\
  ic[xno] \GETS ic[xno] + 2 \cdot length(ls) \\
  \FOREACH nls \in neighbors(ls) \DO
  \BEGIN
    \COMMENT{straight and diagonal offsets of the neighbors} \\
    no \GETS offset(nls) \\
    dno \GETS offset(nls) \\
    dno[0] \GETS 1 \\
    \IF isEmpty(nls) \THEN
    \BEGIN
      \COMMENT{all the lines are on a contour} \\
      \FOREACH l \in ls \DO
      \BEGIN
	ic[no] \GETS ic[no] + length(l)\\
	ic[dno] \GETS ic[dno] + 2 \cdot length(l)
      \END
    \END
    \ELSE
    \BEGIN
      \COMMENT{iterate over the lines in $ls$ and the empty lines in $nls$} \\
      il \GETS iterator(ls) \\
      inl \GETS iterator(nls) \\
      nMin \GETS -\infty \\
      nMax \GETS pos(inl) - 1 \\
      \WHILE isNotAtEnd(il) \DO
      \BEGIN
        lMin \GETS pos(il) \\
        lMax \GETS pos(il) + length(il) - 1 \\
        \COMMENT{measure the intersection of these two lines} \\
	ic[no] \GETS ic[no] + max( 0, min( lMax, nMax ) - max( lMin, nMin ) + 1 ) \\
	ic[dno] \GETS ic[dno] + max( 0, min( lMax, nMax+1 ) - max( lMin, nMin+1 ) + 1 )\\
	ic[dno] \GETS ic[dno] + max( 0, min( lMax, nMax-1 ) - max( lMin, nMin-1 ) + 1 )\\
        \COMMENT{and move to the next line, either in $ls$ or in $nls$} \\
	\IF nMax \leq lMax \THEN
	\BEGIN
          nMin \GETS pos(inl) + length(inl) \\
          next(inl) \\
          \IF isNotAtEnd(inl) \DO
            nMax \GETS pos(inl) - 1
          \ELSE
            nMax \GETS \infty
	\END
	\ELSE
          next(il)
      \END
    \END
  \END
\END
\end{pseudocode}
\caption{\label{interceptCountAlgo}The algorithm used to count the intercepts}
\end{figure}

@@

The roundness is defined as the ratio of equivalent $d$-perimeter over the measured $d$-perimeter. 
The equivalent $d$-perimeter is defined as the perimeter of the $d$-ball with same volume $V$.
\begin{eqnarray}
roundness             & = & \frac{P_{eq}}{P} \\
P_{eq}                & = & \frac{d \cdot V_{eq}}{R_{eq}} \\
R_{eq}                & = & \sqrt[\displaystyle d]{\frac{V\cdot\Gamma(\frac{d+1}{2})}{\pi^\frac{d}{2}}} \\
\Gamma(\frac{d+1}{2}) & = & \begin{cases}
                               \displaystyle \frac{d}{2}!  &\text{if}~ d ~\text{is even} \\
                               \displaystyle \frac{\sqrt{\pi} \cdot d!!}{2^{\frac{d+1}{2}}}  &\text{if}~ d ~\text{is odd} \\
                            \end{cases} \\
d!!                   & = & \begin{cases}
                                1 &\text{if}~ d \leq 1 \\
                                d \cdot (d-2)!!  & \text{otherwise} \\
                             \end{cases}
\end{eqnarray}

@@

Shapes were discretized following the Gauss discretization scheme \cite{Klette2004, Legland2007}.
Binary images can be considered as a subset of a rectangular grid $L_d$ where
d = 2,3. Such a grid can be written as
\begin{eqnarray}
L_d = \Delta_{1} \mathbb{Z} \times \ldots \times \Delta_{1} \mathbb{Z},
\end{eqnarray}
where the $d$-uple $(\Delta_{1},\ldots,\Delta_{d})$ defines the pixel
or voxel size. 
A grid point $x$ belongs to the reconstructed structure if the grid
cell centered on $x$ hits $X$ (See Fig.~\ref{fig:DiskDiscretization}).
